---
title: "Midterm 1 W24"
author: "Clive Smith"
date: "`r Sys.Date()`"
output:
  html_document: 
    keep_md: yes
  pdf_document: default
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your code must be organized, clean, and run free from errors. Remember, you must remove the `#` for any included code chunks to run. Be sure to add your name to the author header above. 

Your code must knit in order to be considered. If you are stuck and cannot answer a question, then comment out your code and knit the document. You may use your notes, labs, and homework to help you complete this exam. Do not use any other resources- including AI assistance.  

Don't forget to answer any questions that are asked in the prompt!  

Be sure to push your completed midterm to your repository. This exam is worth 30 points.  

## Background
In the data folder, you will find data related to a study on wolf mortality collected by the National Park Service. You should start by reading the `README_NPSwolfdata.pdf` file. This will provide an abstract of the study and an explanation of variables.  

The data are from: Cassidy, Kira et al. (2022). Gray wolf packs and human-caused wolf mortality. [Dryad](https://doi.org/10.5061/dryad.mkkwh713f). 

## Load the libraries
```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("janitor")
```

## Load the wolves data
In these data, the authors used `NULL` to represent missing values. I am correcting this for you below and using `janitor` to clean the column names.
```{r message=FALSE, warning=FALSE}
wolves <- read.csv("data/NPS_wolfmortalitydata.csv", na = c("NULL")) %>% clean_names()
```

## Questions
Problem 1. (1 point) Let's start with some data exploration. What are the variable (column) names?  

```{r}
names(wolves)
```

Problem 2. (1 point) Use the function of your choice to summarize the data and get an idea of its structure.  

```{r}
str(wolves)
```

Problem 3. (3 points) Which parks/ reserves are represented in the data? Don't just use the abstract, pull this information from the data.  

```{r}
table(wolves$park)
```
Denali National Park, Grand Teton National Park, Voyageurs National Park, Yellowstone National Park, and Yukon-Charley Rivers National Preserve are represented in the data.

Problem 4. (4 points) Which park has the largest number of wolf packs?

```{r}
wolves %>% 
  group_by(park) %>% 
  summarize(wolf_packs=n_distinct(pack)) %>% 
  arrange(desc(wolf_packs))
```
Denali National Park has the greatest number of wolf packs. 

Problem 5. (4 points) Which park has the highest total number of human-caused mortalities `mort_all`?

```{r}
wolves %>% 
  group_by(park) %>% 
  summarize(h_caused_deaths=sum(mort_all)) %>% 
  arrange(desc(h_caused_deaths))
```
Yukon-Charley Rivers National Preserve has the most human-caused mortalities. 

The wolves in [Yellowstone National Park](https://www.nps.gov/yell/learn/nature/wolf-restoration.htm) are an incredible conservation success story. Let's focus our attention on this park.  

Problem 6. (2 points) Create a new object "ynp" that only includes the data from Yellowstone National Park.  

```{r}
ynp <- wolves %>% 
  filter(park=="YNP")
```


Problem 7. (3 points) Among the Yellowstone wolf packs, the [Druid Peak Pack](https://www.pbs.org/wnet/nature/in-the-valley-of-the-wolves-the-druid-wolf-pack-story/209/) is one of most famous. What was the average pack size of this pack for the years represented in the data?

```{r}
ynp %>% 
  filter(pack=="druid") %>% 
  summarize(mean_pack_size=mean(packsize_aug), 
            n_years=n())
```

Problem 8. (4 points) Pack dynamics can be hard to predict- even for strong packs like the Druid Peak pack. At which year did the Druid Peak pack have the largest pack size? What do you think happened in 2010?

```{r}
ynp %>% 
  filter(pack=="druid") %>% 
   group_by(biolyr) %>% 
  summarize(pack_size=sum(packsize_aug)) %>% 
  arrange(desc(pack_size))
```
In 2010, the pack size was 0. 2010 is also the last year that the Druid Peak pack appears in the data frame. This means that the remaining members of the pack likely died in 2010.

Problem 9. (5 points) Among the YNP wolf packs, which one has had the highest overall persistence `persisty1` for the years represented in the data? Look this pack up online and tell me what is unique about its behavior- specifically, what prey animals does this pack specialize on?  

```{r}
ynp %>% 
  group_by(pack) %>% 
  summarize(pack_persistence=sum(persisty1, na.rm=T)) %>% 
  arrange(desc(pack_persistence))
```
Mollie's Pack has survived with at least 2 members for 26 years, making it the pack with the highest overall persistence in Yellowstone National Park. According to the [Yellowstone Wolf website](https://www.yellowstonewolf.org/yellowstones_wolves.php?pack_id=6), this pack mostly relies on hunting bison to survive. 

Problem 10. (3 points) Perform one analysis or exploration of your choice on the `wolves` data. Your answer needs to include at least two lines of code and not be a summary function.  

I wanted to see which national park wolves had the highest persistence in.

```{r}
wolves %>% 
   group_by(park) %>% 
  summarize(overall_park_persistence=sum(persisty1, na.rm=T), 
            n=n()) %>% 
  arrange(desc(overall_park_persistence))
```
This order makes sense to me, given that Denali also has the highest number of distinct wolf packs. Thinking more logically, this makes sense to me, given that Denali National Park is very large, and is in Alaska, a very remote state. 
