---
title: "BIS15L_AI_hmwk"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
This homework is a bit less traditional this week. There are some errors in the code below. In addition add notes and titles to the code chunks that explains what the code chunk does. Use AI tools, notes from class, or any other resources to help fix the errors. Feel free to edit any plots to your preference. Don't forget to answer the two questions at the bottom. Upon completion, you should not have to '#' out any of the code chunks.

Push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.   

Again make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  


```{r}
library(tidyverse)
library(janitor)
library(dplyr)
```

1.
```{r}
blood_count <- read_csv("data/blood_count_dataset.csv")
```

2. 
```{r}
glimpse(blood_count)
```

3. 

```{r}
blood_count <- clean_names(blood_count)
```

This code chunk gives the min, mean, and max of red blood cell count, grouped by gender.
```{r}
blood_count %>%
  group_by(gender) %>%
  summarize(
    min = min(red_blood_cells),
    mean = mean(red_blood_cells),
    max = max(red_blood_cells)
  )
```

4. 
```{r}
options(scipen = 999, digits = 2)
blood_count %>% 
  filter(gender=="Male") %>%
  select(3:9) %>%
  colMeans(na.rm = TRUE)

```

5. 
```{r}
blood_count %>% 
  filter(gender == "Female") %>%
  select(3:9) %>%
  colMeans(na.rm = T)
```


6. 
```{r}
blood_count %>%
  pivot_longer(cols = [3:9], names_to = "Test", values_to = "Values") %>%
  ggplot(aes(x = Gender, y = Values, fill = Gender)) +
  geom_boxplot() +
  facet(~ Test, scales = "free") 
  labs(title = "Boxplot of Blood Count Tests",
       x = "Test",
       y = "Values")
```

```{r}
blood_count %>%
  pivot_longer(cols = 3:9, names_to = "Test", values_to = "values") %>%
  ggplot(aes(x = gender, y = values, fill = gender)) +
  geom_boxplot() +
  facet_wrap(~ Test, scales = "free") +
  labs(title = "Boxplot of Blood Count Tests",
       x = "Test",
       y = "Values")
```

7. 


```{r}
grouped_bc <- blood_count %>%
  mutate(Age_Group = case_when(
    age >= 20 & age <= 29 ~ "20s",
    age >= 30 & age <= 39 ~ "30s",
    age >= 40 & age <= 49 ~ "40s",
    age >= 50 & age <= 59 ~ "50s",
    age >= 60 ~ "60s")) # Combine 70s with 60s

head(grouped_bc)
```

8. 
```{r}
grouped_bc %>%
  pivot_longer(cols = 3:9, names_to = "Test", values_to = "Values") %>%
  ggplot(aes(x = Age_Group, y = Values, fill = Age_Group)) +
  geom_boxplot() +
  facet_wrap(~Test, scales = "free") +  # Facet by the "Test" variable
  labs(title = "Boxplot of Blood Count Tests by Age Group",
       x = "Age Group",
       y = "Values")
```

9. What are two drawbacks and two benefits of AI? 
This response was generated by chatGPT.
Two Drawbacks of AI:

Job Displacement: One major drawback of AI is the potential for job displacement. As AI systems become more advanced and capable of performing tasks traditionally done by humans, there is a risk of automation leading to job losses in certain industries. This can create economic and social challenges as individuals may need to acquire new skills to remain employable.

Ethical Concerns and Bias: AI systems are only as unbiased as the data they are trained on. If the training data contains biases, the AI system can perpetuate and even exacerbate those biases. Additionally, ethical concerns arise when AI systems are used in sensitive areas such as criminal justice or healthcare, where decisions can have significant consequences. Ensuring fairness, transparency, and accountability in AI systems is an ongoing challenge.

Two Benefits of AI:

Increased Efficiency and Productivity: AI has the potential to significantly enhance efficiency and productivity across various industries. Automation of routine and repetitive tasks allows human workers to focus on more complex and creative aspects of their jobs. AI-powered tools and systems can analyze large datasets, identify patterns, and make predictions at a speed and scale that would be impractical for humans.

Innovation and Problem Solving: AI technologies have the capacity to drive innovation and solve complex problems. They can be applied to fields such as healthcare, finance, and scientific research to discover patterns, optimize processes, and find solutions that may be challenging for humans alone. AI's ability to process vast amounts of data quickly can lead to breakthroughs in areas like drug discovery, climate modeling, and personalized medicine.




10. Do you think you will use AI for coding in the future? Why or why not?

I came up with this response myself.
I will definitely use AI for coding in the future. I think that when paired with a good programmer, AI can make coding more efficient, and help with overcoming roadblocks. I

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences. 