---
title: "BIS15L_AI_class"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Today, we're exploring how AI can aid in writing, comprehending and troubleshooting our R code. AI is a very valuable tool, but it's important not to rely on it entirely. Building a strong R foundation is still important. These AI tools are not perfect and can be inaccurate solutions. In addition these tools can't read our minds, so understanding R helps us generate better prompts to get better responses. In short, AI should be used as a tool, not a crutch. Additionally, do not forget about other useful resources like Stack Overflow/Exchange, online tutorials, and GitHub issues pages for learning and troubleshooting. Let's combine AI with traditional methods to master R effectively.

This overview isn't exhaustive of all possible AI tools for R coding, but I believe it is a good start and covers everything you need.


**Make sure to enable Github Copilot in RStudio.**
Go to Tools -> Global Options -> Copilot -> Enable Git

## Learning Goals: 
*At the end of this exercise, you will be able to:*  
1. Explain what AI is and what AI tools can be used for R.
2. Enable Github Copilot in RStudio.
3. Use Copilot/ChatGPT to write code in RStudio.
4. Use Copilot/ChatGPT for troubleshooting code in RStudio.


## Exploration Questions:
Lets get some get a feel for what copilot (or ChatGPT) does, use Github Copilot to autocomplete the questions below. You can alternatively paste the questions in ChatGPT or Microsoft Copilot.
1. What is AI?

2. What is ChatGPT?

3. What is Github Copilot?

4. [Try it yourself: Insert your own question here]

## Load the libraries  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(dplyr)
library(shiny)
library(shinydashboard)
```

## Load the Data: 
```{r}
#load the Netflix Data in your Data folder, save it under the object name netflix

read_csv("Data/netflix_titles.csv") -> netflix
```

*Did copilot help you write the code above? Could anything about the prompt be improved to help Copilot understand what you want to do more accurately?*


## Lets Practice:
Explore the elements in your data: include a comment within your code chunk to prompt Copilot to use the function of your choice
```{r}
#Use the head function to get the first 6 rows of the netflix data
head(netflix)
```

```{r}
#Use the summary function to get a summary of the netflix data
summary(netflix)
```
```{r}
#use names to get the column names of the netflix data
names(netflix)
```

*Did copilot use your preferred method to get the summary of the data? If not, how could you improve the prompt to get the desired output?*


Sort out the movies from the netflix into an new object called netflix_movies
```{r} 
#use filter to get the movies from the netflix data
filter(netflix, type == "Movie") -> netflix_movies
```


Which director has the most movies on netflix?
```{r}
#use count to get the number of movies by director
count(netflix_movies, director, sort = TRUE, na.rm = TRUE)

```
*Do you see anything wrong with the table above, look at row 1. Write a prompt below to fix this.*

Filter to sort out the director with the most movies but remove the NA's
```{r}
#Filter to sort out the director with the most movies but remove the NA's
netflix_movies %>%
filter(director != "NA") %>% 
  count(director, sort = TRUE)
```


*See anything else interesting about this table, look at the first director on the table. Could this be a problem and how can we edit the dataset to fix this? Try writing a prompt.*

```{r}
netflix_movies %>% 
  filter(director=="Rajiv Chilaka") %>% 
  count(director)
```

Separate each director into their own row

```{r}
#Separate each director into their own row
netflix_movies %>%
  separate_rows(director, sep = ",") %>%
  filter(director != "NA") %>%
  count(director) %>% 
  arrange(desc(n))
```


Make a bar plot of the top directors with the most movies on netflix
```{r}
#Make a bar plot of the top directors with the most movies on netflix
netflix_movies %>%
  separate_rows(director, sep = ",") %>%
  filter(director != "NA") %>%
  count(director) %>% 
  arrange(desc(n)) %>% 
  top_n(10) %>% 
  ggplot(aes(x = reorder(director, n), y = n)) +
  geom_bar(stat = "identity", fill="dodgerblue2", color="black", alpha=0.4) +
  theme_linedraw() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(title = "Top 10 Directors with the most movies on Netflix",
       x = "Director",
       y = "Number of Movies")
```
*Do you see any issues with this plot? How could we improve the prompt to get a better boxplot? Try writing a prompt.*

[Insert your prompt here]
```{r}

```

*How about for just 2020*



Make a bar plot of the top directors with the most movies on netflix in 2020
```{r}
#Make a bar plot of the top directors with the most movies on netflix in 2020
netflix_movies %>%
  separate_rows(director, sep = ",") %>%
  filter(director != "NA") %>%
  filter(release_year == 2020) %>%
  count(director) %>% 
  arrange(desc(n)) %>% 
  top_n(10) %>% 
  ggplot(aes(x = reorder(director, n), y = n)) +
  geom_bar(stat = "identity", fill="dodgerblue2", color="black", alpha=0.4) +
  theme_linedraw() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(title = "Top 10 Directors with the most movies on Netflix in 2020",
       x = "Director",
       y = "Number of Movies")
```


*Take a few minutes to make your own plot(s) using prompts, edit prompts until you get a plot that you like.*


## Lets make the top 10 directors barplot into a Shiny App: use copilot or Chatgpt

```{r}
#Make the top 10 directors barplot into a Shiny App

ui <- fluidPage(
  titlePanel("Top 10 Directors with the most movies on Netflix"),
  sidebarLayout(
    sidebarPanel(
      sliderInput("year", "Year", min = 2010, max = 2020, value = 2020)
    ),
    mainPanel(
      plotOutput("top_directors")
    )
  )
)



server <- function(input, output) {
  output$top_directors <- renderPlot({
    netflix_movies %>%
      separate_rows(director, sep = ",") %>%
      filter(director != "NA") %>%
      filter(release_year == input$year) %>%
      count(director) %>% 
      arrange(desc(n)) %>% 
      top_n(10) %>% 
      ggplot(aes(x = reorder(director, n), y = n)) +
      geom_bar(stat = "identity", fill="dodgerblue2", color="black", alpha=0.4) +
      theme_linedraw() +
      theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
      labs(title = "Top 10 Directors with the most movies on Netflix",
           x = "Director",
           y = "Number of Movies")
  })
}

shinyApp(ui = ui, server = server)

```

*Now use ChatGPT, Copilot or previous notes to fix the code*






## That's it! Let's take a break and then move on to working on group projects.
