---
title: "midterm_1_cheatsheet.rmd"
author: "Clive Smith"
date: "2024-02-04"
output: html_document
---

list of useful functions for midterm 1.

filter()
used to look through observations and pull out or exclude values that match a certain condition. Observations must meet both conditions when "&" and "," operators are used, they must only meet one condition when "|" operator is used.
examples:
filter(data_frame, variable=="NAME")
filter(data_frame, variable==number), can also use >, <, >=, <=.
filter(data_frame, variable1=="NAME" | variable2>=number)

select()
used to pull out variables that meet a certain condition. 
select_if(.==TRUE | .=="One Punch Man")
mutate()
used to create a new variable by transforming another variable in some way.
Example:

```{r}
#data_frame %>% 
#mutate(new_variable=old_variable/2)
```


#Lab 2:
spring_1 <- c(36.25, 35.40, 35.30)
spring_2 <- c(35.15, 35.35, 33.35)
spring_3 <- c(30.70, 29.65, 29.20)
spring_4 <- c(39.70, 40.05, 38.65)
spring_5 <- c(31.85, 31.40, 29.30)
spring_6 <- c(30.20, 30.65, 29.75)
spring_7 <- c(32.90, 32.50, 32.80)
spring_8 <- c(36.80, 36.45, 33.15)

spring_temps <- c(spring_1, spring_2, spring_3, spring_4, spring_5, spring_6, spring_7, spring_8)

spring_matrix <- matrix(spring_temps, ncol = 3, byrow=T)

colnames(spring_matrix) <- scientists
rownames(spring_matrix) <- spring_names

rowMeans(spring_matrix)

cbind(spring_matrix, mean_spring_temp)

#Lab 3:
sleep <- data.frame(msleep)
##Important!!! Functions for statistical summaries of data sets
dim(sleep)
table(sleep$vore)
```{r}
#sleep %>% 
  #count(vore)
```
summary(sleep)
names(sleep)
glimpse(sleep)
#Lab 4
homerange <- read_csv("data/Tamburelloetal_HomeRangeDatabase.csv") #make sure working directory is correct for this.
homerange$taxon <- as.factor(homerange$taxon)
levels(homerange$taxon) #shows levels of a variable with class factor.
 #Filter and select can be paired together to restrict the values in a data frame. USE FILTER BEFORE SELECT!!!
```{r}
#owls <- homerange %>% 
  #filter(order=="strigiformes") %>% 
  #select(common.name, mean.mass.g, log10.mass, family, genus, species) %>% 
  #arrange(mean.mass.g)
```
head(owls, 1)

#Lab 5 
clean_names(data_frame)

```{r}
#fish %>% 
  #mutate(length_mm = length*10)
```
```{r}
#mammals %>%
  #mutate_all(tolower)
```
```{r}
#mammals %>%
  #mutate(newborn_new = ifelse(newborn == -999.00, NA, newborn)
```

#Lab 6
tabyl(superhero_info, alignment)

#Lab 7
skim(msleep24)
hist(msleep24$sleep_total_24)

```{r}
#msleep %>% 
  #filter(bodywt>200) %>% 
  #summarize(mean_sleep_lg=mean(sleep_total))
```

```{r}
#msleep %>% 
  #filter(bodywt>200) %>% 
  #summarize(mean_sleep_lg=mean(sleep_total), 
            #min_sleep_lg=min(sleep_total), 
            #max_sleep_lg=max(sleep_total),
            #sd_sleep_lg=sd(sleep_total),
            #total=n())
``` 
summarize(n_genera=n_distinct(genus)) #this is going to count the number of genera in msleep.

group_by(vore) #we are grouping by feeding ecology, a categorical variable.
You can only group by categorical variables. 

summarize(number_NAs=sum(is.na(body_mass_g)))

penguins %>% 
  filter(!is.na(body_mass_g)) %>% #pull out all of the observations with a number.
  
```{r}
#penguins %>% 
  #count(island, sort = T) sort=T sorts the column in descending order
```
  
```{r}
#penguins %>% 
  #count(island, species, sort = T) sort=T will arrange in descending order
```

```{r}
#penguins %>% 
  #tabyl(island, species)
```

```{r}
#penguins %>%
  #group_by(sex) %>%
  #summarize(across(contains("mm"), mean, na.rm=T))
```
summarize_all goes across all variables.
