---
title: "lab_6_warmup.Rmd"
output: html_document
date: "2024-01-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load the libraries
```{r}
library(tidyverse)
library(janitor)
```

##1. Load the `bison.csv` data.
```{r}
bison <- read_csv("data/bison.csv")
```
##2. What are the dimensions and structure of the data?
```{r}
dim(bison)
```

```{r}
summary(bison)
```
##3. We are only interested in code, sex, weight, year of birth. Restrict the data to these variables and store the dataframe as a new object.
##4. Pull out the animals born between 1980-1990.
```{r}
names(bison)
```

```{r}
bison_data <- bison %>% 
  select(contains("animal")) %>% 
  filter(between(animal_yob, 1980, 1990))
bison_data
```

##5. How many male and female bison are represented between 1980-1990?
```{r}
table(bison_data$animal_sex)
```

##6. Between 1980-1990, were males or females larger on average?
```{r}
 females <- filter(bison_data, animal_sex=="F")
mean(females$animal_weight)
```

```{r}
bison_data %>% 
  filter(between(animal_yob, 1980, 1990)) %>% 
  filter(animal_sex=="M")
```

```{r}
males <- filter(bison_data, animal_sex=="M")
mean(males$animal_weight)
```

males born between 1980 and 1990 are larger on average than females born during the same time frame.