---
title: "Midterm Notes"
author: "Clive Smith"
date: "2024-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load library:

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(skimr)
```

>>>
# Lab 2: 

```{r}
object <- matrix(dataset.data, nrow=n)
```


```{r}
object <- cbind(dataset, row_name = object_with_mean) #allows me to addd coloumn
```

>>>
# Lab 3: 

```{r}
object <- data.frame(matrix_1, matrix_2, matrix_2)
```

```{r}
write.csv(dataset, "name_of_it", row.names = FALSE)
object <- read_csv("folder/dataset_name")
library(janitor)
clean_names(dataset)
object <- readr::read_csv("folder/dataset_name")
```


For basic analysis:
```{r}
str(dataset) 
glimpse(dataset)
names(dataset)
summary(dataset)
dim(dataset)
```

```{r}
head(dataset, n) #gives you first row frame
tail(dataset, n) #last rows 

table(dataset$column)
count(dataset$column) #for the pipe
```



Mean alternative:
```{r}
object <- t.test(large$sleep_total, small$sleep_total)
object 
```

Mean compersion: 
```{r}
if (mean(filtered_dataset$column) > mean(filtered_dataset.1$column)) {
  cat("The mean of 'filtered_dataset' is bigger.\n")
} else if (mean(filtered_dataset1$column) > mean(filtered_dataset$column)) {
  cat("The mean of 'filtered_dataset1' is bigger.\n")
} else {
  cat("The means are equal.\n")
}
```



Finding max of ceratin colums:
```{r}
object <- max(dataset$column)
object_2 <- dataset[dataset$column == object, c("coulumn", "column", "column")]
object_2
```

>>>
 Lab 4:

```{r}
select(dataset, matches/contains/ends_with("a.+er")) #handy bit of code for the project (parts of words)

select(dataset, column)

select_if(dataset, ~!is.numeric(.)) # (.) all of the cells, ! - all other then mentions data 

filter(dataset, column %in% c(167, 175))
```

```{r}
select(fish, -"fish_id", -"annnumber", -"length", -"radii_length_mm")
```

```{r}
largest_home_range <- vertebrate_clean %>%
  filter(class == "aves" & !is.na(mean_hra_m2)) %>% #for data where we have NA
  arrange(desc(mean_hra_m2)) %>%
  head(1) %>%
  select(mean_hra_m2, common_name, genus, species)
largest_home_range
```

>>>
# Lab 5

```{r}
filter(dataset, column == "AL" & column1 > 350) 
```


Here we compare means
```{r}
if (mean(mammals$mass[mammals$genus == "Lophocebus"]) > 
    mean(mammals$mass[mammals$genus == "Erythrocebus"]) &&
    mean(mammals$mass[mammals$genus == "Erythrocebus"]) > 
    mean(mammals$mass[mammals$genus == "Macaca"])) {
  cat("Lophocebus has the smallest newborn mass among the specified genera.\n")
} else if (mean(mammals$mass[mammals$genus == "Erythrocebus"]) > 
           mean(mammals$mass[mammals$genus == "Lophocebus"]) &&
           mean(mammals$mass[mammals$genus == "Lophocebus"]) > 
           mean(mammals$mass[mammals$genus == "Macaca"])) {
  cat("Erythrocebus has the smallest newborn mass among the specified genera.\n")
} else if (mean(mammals$mass[mammals$genus == "Macaca"]) > 
           mean(mammals$mass[mammals$genus == "Lophocebus"]) &&
           mean(mammals$mass[mammals$genus == "Lophocebus"]) > 
           mean(mammals$mass[mammals$genus == "Erythrocebus"])) {
  cat("Macaca has the smallest newborn mass among the specified genera.\n")
} else {
  cat("The means are equal.\n")
}
```




```{r}
fish %>% 
  select(lakeid, radii_length_mm) %>% 
  filter(lakeid == "AL" | lakeid == "AR") %>% 
  filter (between(radii_length_mm, 2, 4)) %>% 
  arrange(desc(radii_length_mm)) #here we combine AR and AL and sort them from high to low
```

```{r}
fish %>% 
  mutate(length_mm = length*10) %>% 
  select(fish_id, length, length_mm) # allows to create new column fro existing one
```
  
```{r}
mammals %>%
  mutate_all(tolower) # all to lower case

```


>>>
# Lab 6 

Pullinng out top value: 
```{r}
mammals %>% 
  select(genus, species, max_life) %>%
  mutate(max_life_new = ifelse(max_life == -999.00, NA, max_life))%>% 
  arrange(desc(max_life)) %>%
  slice(1)
```



Find your hero in the superhero_info data? Show their info!  
```{r}
my_favorite <- superhero_power %>%
  filter(hero_names == "Groot") %>%
  select_if(~all(. == "TRUE") | . == "Groot")
my_favorite
```

```{r}
my_hero_info <- superhero_power %>%
  filter(hero_names %in% my_favorite$hero_names)
my_hero_info
```



Basic data analysis 
```{r}
tabyl(dataset, column)
```

>>>
# Lab 7

```{r}
msleep24$sleep_total_24 <- as.numeric(msleep24$sleep_total_24)
hist(msleep24$sleep_total_24, main = "Histogram of sleept_total", xlab = "sleept_total") 
```




```{r}
msleep %>%
  filter(bodywt > 200) %>%
  summarise(mean_sleep_lg = mean(sleep_total),
            min_sleep_lg = min(sleep_total),
            max_sleep_lg = max(sleep_total), 
            sd_sleep_lg = sd(sleep_total), 
            total=n())
```
            
```{r}
msleep %>% 
  filter(bodywt > 100) %>%
  n_distinct()
```

```{r}
msleep %>%
  group_by(order) %>%
  summarize(min_brainwt = min(brainwt, na.rm=TRUE),
            max_brainwt = max(brainwt, na.rm=TRUE),
            mean_brainwt = mean(brainwt, na.rm=TRUE),
            total=n())
```


            
```{r}
penguins %>%
  group_by(species, island) %>%
  filter(!is.na(body_mass_g)) %>%
 summarize(mean_body_mass_g = mean(body_mass_g))
```

```{r}
penguins %>% 
  group_by(species, island) %>% 
  summarize(n=n(), .groups= 'keep')#the .groups argument here just prevents a warning message, n=n is a number of observations
```

  
  
```{r}
penguins %>% 
  count(island, species, sort = T) # sort=T will arrange in descending order
```

```{r}
penguins %>%
  group_by(species, island) %>%
  summarize(n_nas = sum(is.na(sex)))
```
  
```{r}
penguins %>%
  summarize(across(contains("mm"), \(x) mean(x, na.rm = TRUE))) #to get rid of R console, (\(x))
```
  
```{r}
penguins %>%
  summarise(across(starts_with("bill"), mean, na.rm=T))
```

## ggplot aesthetics

```{r}
  geom_picksomething(color="black", fill="dodgerblue2", alpha=0.4)+
  theme_linedraw()+
  theme(plot.title=element_text(size = rel(1.75), hjus=0.5), axis.text.x = element_text(angle = 60,     hjust=1))+
  labs(title="Life Expectancy by Year", 
       x=NULL, y="Life Expectancy in Years")
```
  
##lab 8

For a summary of NAs in the data.
```{r}
life_history %>% 
  map_df(~ sum(is.na(.)))
```

One way to manage NAs (not very efficient).
```{r}
life_history %>% 
  mutate(wean_mass_new=ifelse(wean_mass==-999, NA, wean_mass))
```

Easiest way to manage NAs, but also can be quite risky.
```{r}
life_history <- life_history <- read_csv("data/mammal_lifehistories_v3.csv", na= c("NA", " ", ".", "-999", "not measured")) %>% 
  clean_names()
# you need to know how the NA's are represented in the data
# you don't want to do this by default
```

better way to manage NAs.
```{r}
amniota_tidy <- amniota %>% 
  replace_with_na_all(condition = ~.x == -999)
```

Another way to manage NAs; more targeted.
```{r}
life_history %>% 
  replace_with_na(replace = list(newborn = "not measured", 
                                 weaning= -999, 
                                 wean_mass= -999, 
                                 afr= -999, 
                                 max_life= 0, 
                                 litter_size= -999, 
                                 gestation= -999, 
                                 mass= -999))
```

summarizes NAs (assuming they're not classified as something else).
```{r}
amniota %>% 
  summarize(number_NAs=sum(is.na(amniota)))
```

another way to summarize NAs.
```{r}
miss_var_summary(amniota_tidy)
```

##Lab 9

Use pivot longer as necessary to make tidy data.

pivot longer.
```{r}
heartrate %>% 
  pivot_longer(-patient, #patient will not move
               names_to = "drug", #make a new column called "drug"
               values_to="heartrate" #values moved to a new column called "heartrate"
               )
```

```{r}
billboard %>% 
   pivot_longer(
   cols = starts_with("wk"), #columns that start with "wk"
   names_to = "week",
   names_prefix = "wk",
   values_to = "rank",
   values_drop_na = TRUE)
```

separate.
```{r}
heartrate2 %>% 
  separate(patient, into= c("patient", "sex"), sep = "_")
```

unite
```{r}
heartrate3 %>% 
  unite(patient_sex, "patient", "sex", sep=" ")
```

pivot wider
`values_from` to identify the values associated with the new columns.
```{r}
tb_data %>% 
  pivot_wider(names_from = "key", #the observations under key will become new columns
              values_from = "value") #the values under value will be moved to the new columns
```

##lab 10

makes a boxplot
```{r}
homerange %>% 
  ggplot(aes(x = class, y = log10.mass)) +
  geom_boxplot()
```

Makes a bar plot (use stat="identity" inside of geom_bar() to specify 2 variables)
```{r}
homerange %>% 
  filter(family=="salmonidae") %>%
  select(common.name, log10.mass) %>% 
  ggplot(aes(x=common.name, y=log10.mass))+ #notice the switch in x and y
  geom_col()+
  coord_flip()
```

##lab 11

```{r}
options(scipen=999)#cancels scientific notation for the session
```

```{r}
scale_y_log10() #scales y-axis by log_10.
```

x=(reorder(order, mean_mass)) reorders the x-axis, order, by mean_mass values.
```{r}
life_history %>% 
  group_by(order) %>% 
  summarize(mean_mass=mean(mass, na.rm=T)) %>% 
  ggplot(aes(x=reorder(order, mean_mass), y=mean_mass))+ #this allows us to order the x axis by mean_mass
  geom_col()
```

Makes a scatterplot (geom_jitter reduces overplotting)
```{r}
life_history %>% 
  ggplot(aes(x=gestation, y=wean_mass))+
  geom_jitter(na.rm=T)+ #prevents overplotting
  scale_y_log10()
```

Overlays a line of best fit onto a scatterplot.
```{r}
elephants %>% 
  ggplot(aes(x=age, y=height))+
  geom_point()+
  geom_smooth(method=lm, se=F)
```

`size` adjusts the size of points relative to a continuous variable.
```{r}
life_history %>% 
  ggplot(aes(x=gestation, y=log10(mass), size=mass))+
  geom_point(na.rm=T)
```

We can also map shapes to another categorical variable.
```{r}
p + geom_point(aes(color=thermoregulation, shape=thermoregulation), size=1.5)
```

We can also scale all bars to a percentage.
```{r}
homerange %>% 
  ggplot(aes(x = taxon, fill = trophic.guild))+
  geom_bar(position = position_fill())+ 
  scale_y_continuous(labels = scales::percent)+
  coord_flip()
```

position="dodge" puts things onto different columns.
```{r}
position="dodge"
```

##lab 12
line plot
```{r}
deserts %>% 
  filter(species_id=="DM"| species_id=="DS") %>% 
    mutate(year=as.factor(year)) %>% 
  group_by(year, species_id) %>% 
  summarize(n=n(), .groups='keep') %>% 
  ggplot(aes(x=year, y=n, group=species_id, color=species_id))+
  geom_line()+
  geom_point(shape=1)+ #you can experiment with shapes
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Number of Samples for Species DM & DS", 
       x="Year", y="n")
```

```{r}
homerange %>% 
  ggplot(aes(x = log10.mass)) +
  geom_histogram(bins = 20)+ #we can adjust the number of bins with the bins argument
  labs(title = "Distribution of Body Mass")
```

makes a curve (similar to a histogram, but for continuous variables)
```{r}
homerange %>% 
  ggplot(aes(x=log10.hra))+
  geom_density(fill = "violetred4", alpha = 0.4, color = "black")+
  labs(title = "Distribution of Homerange")
```

case_when
```{r}
homerange %>% 
  mutate(range_category=case_when(log10.hra<3.65 ~ "very small", 
                                  log10.hra>=3.65 & log10.hra<=4.59 ~ "small", 
                                  log10.hra>4.59 & log10.hra<=6.02 ~ "medium", 
                                  log10.hra>6.02 ~ "large"))
```

facet_wrap()
```{r}
homerange %>% 
  ggplot(aes(x=taxon, y=log10.mass))+
  geom_boxplot()+
  facet_wrap(~trophic.guild)+
  theme(axis.text.x = element_text(angle = 60, hjust=1))
```

`facet_grid()` will also allow the comparison of two categorical variables, just remember a~b where a is rows and b is columns.  
```{r}
homerange %>% 
  ggplot(aes(x=taxon, y=log10.mass))+
  geom_boxplot()+
  facet_grid(trophic.guild~thermoregulation)+
  theme(axis.text.x = element_text(angle = 60, hjust=1))
```
